---
title: "Excerpt from Mid-Atlantic Ecosystem Approach"
subtitle: "Overview and Progress <br /> May 2020"
author: "Sarah Gaichas, Geret DePiper, Brandon Muffley, Richard Seagraves <br /> Northeast Fisheries Science Center, Mid-Atlantic Fishery Management Council"
output:
  xaringan::moon_reader:
    seal: false
    css: ["default", "libs/EDAB_theme3_508_SOE.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      titleSlideClass: ["right", "middle", "my-title"]
      highlightStyle: githubp
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "libs/macros.js"
  revealjs::revealjs_presentation:
    self_contained: false
    reveal_plugins: ["notes", "search"]
---
class: top, left

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = F,
                      warning = F,
                      message = F)
#Plotting and data libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(here)
library(kableExtra)
library(ggrepel)
library(stringr)
library(patchwork)
library(grid)
library(plotly)
library(vegan)
library(rpart)
library(colorRamps)
library(ecodata)

#GIS libraries
library(sf)
library(rgdal)
#library(raster)
library(rnaturalearth)

data.dir <- here::here("data")


#Time series constants
shade.alpha <- 0.3
shade.fill <- "lightgrey"
lwd <- 1
pcex <- 2
trend.alpha <- 0.5
trend.size <- 2
hline.size <- 1
hline.alpha <- 0.35
hline.lty <- "dashed"
label.size <- 5
hjust.label <- 1.5
letter_size <- 4
feeding.guilds <- c("Apex Predator","Piscivore","Planktivore","Benthivore","Benthos")
x.shade.min <- 2009
x.shade.max <- 2019
map.lwd <- 0.4
#Function for custom ggplot facet labels
label <- function(variable,value){
  return(facet_names[value])
}

#facet names for titles
facet_names <- list("Apex predators" = expression("Apex predators"),
                    "Piscivores" = expression("Piscivores"),
                    "Planktivores" = expression("Planktivores"),
                    "Benthivores" = expression("Benthivores"),
                    "Benthos" = expression("Benthos"))
#CRS
crs <- "+proj=longlat +lat_1=35 +lat_2=45 +lat_0=40 +lon_0=-77 +x_0=0 +y_0=0 +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"

#Coastline shapefile
coast <- ne_countries(scale = 10,
                          continent = "North America",
                          returnclass = "sf") %>%
             sf::st_transform(crs = crs)

#State polygons
ne_states <- ne_states(country = "united states of america",
                                      returnclass = "sf") %>%
  sf::st_transform(crs = crs)

#high-res polygon of Maine
#new_england <- read_sf(gis.dir,"new_england")

#EPU shapefile
epu_sf <- ecodata::epu_sf %>% 
  filter(EPU %in% c("MAB","GB","GOM"))

#Region identifiers
epu <- "Mid-Atlantic Bight"
epu_abbr <- "MAB"
region <- "Mid-Atlantic"
region_abbr <- "MA" #Some commercial data organized by "MA" or "NE" regions, not by EPU 


```

## Risk assessent indicators and ranking criteria: Commercial revenue 

This element is applied at the ecosystem level. Revenue serves as a proxy for commercial profits.

.table[
.pull-left[

```{r riskcomval, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| Risk Level         | Definition                                                          |  
|:-------------------|:--------------------------------------------------------------------|
| Low  | No trend and low variability in revenue |
| Low-Moderate | Increasing or high variability in revenue |
| Moderate-High | Significant long term revenue decrease |
| High | Significant recent decrease in revenue |
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```
]
]
.pull-right[
```{r total-rev, fig.width = 6, fig.asp = 0.45, fig.retina=3, fig.cap = "Total revenue for the region (black) and revenue from MAFMC managed species (red)."}
#Filtering and aggregation step
rev_agg <- ecodata::comdat %>% 
  filter(str_detect(Var, "Revenue"),
         !str_detect(Var, "prop|Other|NEFMC"), #Remove proportions, "Other" category species, NEFMC managed species in MAB
         EPU == epu_abbr,
         Time >= 1986) %>% 
  mutate(Status = ifelse(str_detect(Var, "Revenue weight"), 
                         "Managed","Total")) %>% #Create groups for aggregation
  group_by(Status, Time) %>% 
  dplyr::summarise(Total = sum(Value)) %>% 
  group_by(Status) %>% 
  mutate(hline = mean(Total))

series.col <- c("indianred","black")

#Plotting
ggplot(data = rev_agg) +
  
  #Highlight last ten years
  annotate("rect", fill = shade.fill, alpha = shade.alpha,
      xmin = x.shade.min , xmax = x.shade.max,
      ymin = -Inf, ymax = Inf)+  
  
  #lines
  geom_gls(aes(x = Time, y = Total,
               group = Status),
             alpha = trend.alpha, size = trend.size) +
  geom_line(aes(x = Time, y = Total, color = Status), size = lwd) +
  geom_point(aes(x = Time, y = Total, color = Status), size = pcex) +

  #axes
  scale_y_continuous(labels = function(l){trans = l / 1000000})+
  scale_x_continuous(breaks = seq(1985, 2015, by = 5), expand = c(0.01, 0.01)) +
      scale_color_manual(values = series.col, aesthetics = "color")+
  guides(color = FALSE) +
  ggtitle("Total Revenue") +
  ylab(expression("Revenue (10"^6*"USD)")) +
  geom_hline(aes(yintercept = hline,
               color = Status),
           size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty) +
  theme_ts()
```

]

Ranked moderate-high risk: significant long term decrease in Mid-Atlantic managed species revenue (red points)

*All ecosystem level risk element rankings*
```{r ecotable}
#tab.cap="Ecosystem level risk analysis results; l=low risk (green), lm= low-moderate risk (yellow), mh=moderate to high risk (orange), h=high risk (red)\\label{sptable}",

risk.eco<-data.frame(
  System = c("Mid-Atlantic"),
  EcoProd = c("lm"),
  #EcoDiv = c("lm"),
  CommRev = c("mh"),
  RecVal = c("h"),
  FishRes1 = c("l"),
  FishRes4 = c("mh"),
  #CommJobs = c("mh"),
  #RecJobs = c("l"),
  FleetDiv = c("l"),
  Social = c("lm"),
  ComFood = c("h"),
  RecFood = c("mh")
)

#convert to longer text for consistency and display in html table
risk.eco <- risk.eco %>%
     mutate_at(vars(-System), function(x){
       recode(x,'l'="lowest",'lm'="lowmod",'mh'="modhigh",'h'="highest")}) %>%
     as.data.frame()

#make table
risk.eco %>%
  mutate_at(vars(-System), function(x){ 
    cell_spec(x, format="html", color = factor(x, c("na", "lowest", "lowmod", "modhigh", "highest"),c("black", "black", "black", "black", "white")), align = "c", background =factor(x, c("na", "lowest", "lowmod", "modhigh", "highest"),c("white", "lightgreen", "yellow", "orange", "red")), background_as_tile=F)}) %>%
  kable(format = "html", escape = F, table.attr='class="risktable"') %>%
  kable_styling(font_size = 14) 

```
